\huge Gap formation and stability in non-isothermal protoplanetary disks \large [2014 CITA Summer Student Program]

### 2014-07-24 poster/poster.tex
minor poster changes

\title{\huge Gap formation and stability in non-isothermal
  disks\\ \Large [2014 CITA Summer Student Program]}
\author{Min-Kai Lin and Robert Les$^\dagger$}
\institute[CITA]{Canadian Institute for Theoretical Astrophysics, 60
  St George Street, Toronto, M5S 3H8, Canada\\{\footnotesize$^\dagger$student: all
  simulations, data analysis and figures}}

@@

            \begin{block}{{\Large Disk-planet model with simple
                  cooling}}
              \justifying
             We numerically evolve the 2D fluid equations including a
             gap-opening giant planet, treated as an external
             potential on fixed circular orbit. We include in the
             energy equation  
        a simple source term in the form 
	\begin{align*}
	\frac{\partial e}{\partial t} = \cdots + \frac{e-e_{t=0}}{t_\mathrm{cool}},
	\end{align*}
        and $t_\mathrm{cool} = \beta \Omega_k^{-1}$, where $\Omega_k$
        is the Keplerian frequency. The energy density
        $e=p/(\gamma-1)$ where $p$ is the pressure field and
        $\gamma=1.4$ is the adiabatic index. We choose $\beta$ indirectly 
        through $\tilde{\beta}$ such that
        \begin{align*}
        t_\mathrm{cool, gap\,edge} = \tilde{\beta} t_\mathrm{lib, gap\,edge},
        \end{align*}
\vspace{-1.8cm}
        \begin{itemize}
        \item  $t_\mathrm{lib, gap\,edge}$: time interval between
          successive encounters of a fluid element located at the gap
          edge, and the planet's azimuth
        \item $\tilde{\beta}\ll 1 \Rightarrow$ fluid at the gap edge returns to its initial temperature
          between going through planet-induced shocks 
        \end{itemize}
        \end{block}
        \vfill

@@

          density perturbations. Fourier analysis of the density field
          gives the dominant mode number $m$ and growth rate $\gamma$.  

@@

            \begin{block}{\Large{Non-linear evolution of gap edge
                  vortices}}
              \justifying
%                  {\bf `lifetime' plot of m=1 for diff beta, snapshots
%                    (before, during, after vortex decay). quote rossby
%                  number when vortices in quasi-steady state}
              Generic evolution independent of $t_\mathrm{cool}$:
              \begin{itemize}
              \item Single vortex in quasi-steady state from vortex merging
              \item Vorticity source from disk-planet interaction
                $\to$ vortex intensifies
              \item Vortex dissipates quickly after its Rossby number
                $Ro$ minimizes, due to shocks and/or smoothing out the
                gap edge  
              \end{itemize}
                  \begin{figure}
                    \centering
                    \hfill
                    \begin{minipage}{0.3\textwidth}
                      \includegraphics[width=\textwidth]{Posterfig_Before}
                      \caption{$Ro=-0.16$}
                    \end{minipage}
                    \hfill
                    \begin{minipage}{0.3\textwidth}
                      \includegraphics[width=\textwidth]{Posterfig_During}
                      \caption{$Ro=-0.43$}
                    \end{minipage}
                    \hfill
                    \begin{minipage}{0.3\textwidth}
                      \includegraphics[width=\textwidth]{Posterfig_After}
                      \caption{$Ro=-0.10$}
                    \end{minipage}
                    \hfill
                  \end{figure}

@@

            \begin{block}{\Large{Vortex lifetimes and cooling rate}}
              \justifying
              \begin{itemize}
              \item Vortices typically live longer with increasing cooling time:
              \end{itemize}
              \begin{figure}
                \centering
                \includegraphics[width=\textwidth,clip=true,trim=0cm
                  1.5cm 0cm 1cm]{lifetime_regime1}
              \end{figure}
              \vspace{-1cm}
              \begin{itemize}
              \item However, for very long $t_\mathrm{cool}$ vortex
                lifetimes shorten again:
              \end{itemize}
              \begin{figure}
                \centering
                \includegraphics[width=\textwidth,clip=true,trim=0cm
                  0cm 0cm 1cm]{lifetime_regime2}
              \end{figure}
              \vspace{-2cm}
              \begin{itemize}
              \item A cooling time of $\tilde{\beta}\sim 5$ maximizes
                vortex lifetime (nearly adiabatic)
              \end{itemize}
            \end{block}
            \vfill
            \begin{block}{{\Large Summary}}\justifying
              {\large{\bf
                  The stability of planetary gaps in gaseous
                  protoplanetary disks is affected by disk
                  thermodynamics. A slower cooling rate, which permits
                  more heating, tends to stabilize the system, but
                  also enables longer vortex lifetimes unless the
                  cooling rate is extremely long. Our simulations
                  suggest there exists an optimal $t_\mathrm{cool}$
                  that maximizes vortex lifetimes, a result consistent
                  with recent isothermal simulations by Fu et
                  al. (2014). 
                }
              }
              \vfill
            \end{block}
          }
          % ---------------------------------------------------------%
          % end the column
        \end{minipage}
      \end{beamercolorbox}
    \end{column}
    % ---------------------------------------------------------%
    % end the column
  \end{columns}
  \vskip1ex
  %\tiny\hfill\textcolor{ta2gray}{Created with \LaTeX \texttt{beamerposter}  \url{http://www-i6.informatik.rwth-aachen.de/~dreuw/latexbeamerposter.php}}
  %% \tiny\hfill{Created with \LaTeX \texttt{beamerposter}  \url{http://www-i6.informatik.rwth-aachen.de/~dreuw/latexbeamerposter.php} \hskip1em}
\end{frame}
\end{document}

>>>>

\title{\huge Gap formation and stability in non-isothermal
  protoplanetary disks \large [2014 CITA Summer Student Program]}
\author{Min-Kai Lin and Robert Les$^\dagger$}
\institute[CITA]{Canadian Institute for Theoretical Astrophysics, 60
  St George Street, Toronto, M5S 3H8, Canada\\{\footnotesize$^\dagger$student: all
  simulations, data analysis and figures}}

@@

            \begin{block}{{\Large Disk-planet model with simple
                  cooling}}
              \justifying
             We numerically evolve the 2D fluid equations including a
             gap-opening giant planet, treated as an external
             potential on fixed circular orbit. We include in the
             energy equation  
        a simple source term in the form 
	\begin{align*}
	\frac{\partial e}{\partial t} = \cdots + \frac{e-e_{t=0}}{t_\mathrm{cool}},
	\end{align*}
        and $t_\mathrm{cool} = \beta \Omega_k^{-1}$, where $\Omega_k$
        is the Keplerian frequency. The energy density
        $e=p/(\gamma-1)$ where $p$ is the pressure field and
        $\gamma=1.4$ is the adiabatic index. We choose $\beta$ indirectly 
        through $\tilde{\beta}$ such that
        \begin{align*}
        t_\mathrm{cool, gap\,edge} = \tilde{\beta} t_\mathrm{lib, gap\,edge},
        \end{align*}
\vspace{-1.6cm}
        \begin{itemize}
        \item  $t_\mathrm{lib, gap\,edge}=2\pi/|\Omega_k(r_p)-\Omega_k(r_\mathrm{edge})|$: time between
          successive encounters of a fluid element at the gap
          edge and the planet's azimuth
        \item $\tilde{\beta}\ll 1 \Rightarrow$ fluid at the gap edge
          returns to its initial temperature inbetween going through planet-induced shocks 
        \end{itemize}
        \end{block}
        \vfill

@@

          density perturbations. We use Fourier analysis to find
          the dominant mode number $m$ and growth rate $\gamma$.  

@@

            \begin{block}{\Large{Non-linear evolution of gap edge
                  vortices}}
              \justifying
%                  {\bf `lifetime' plot of m=1 for diff beta, snapshots
%                    (before, during, after vortex decay). quote rossby
%                  number when vortices in quasi-steady state}
              Generic evolution independent of $t_\mathrm{cool}$:
              \begin{itemize}
              \item Single vortex in quasi-steady state from vortex merging
              \item Continous vorticity source from disk-planet interaction
                $\to$ vortex intensifies
              \item Vortex dissipates quickly after its Rossby number
                $Ro$ minimizes, due to shocks and/or smoothing out the
                gap edge  
              \end{itemize}
                  \begin{figure}
                    \centering
                    \hfill
                    \begin{minipage}{0.3\textwidth}
                      \includegraphics[width=\textwidth]{Posterfig_Before}
                      \caption{$Ro=-0.16$}
                    \end{minipage}
                    \hfill
                    \begin{minipage}{0.3\textwidth}
                      \includegraphics[width=\textwidth]{Posterfig_During}
                      \caption{$Ro=-0.43$}
                    \end{minipage}
                    \hfill
                    \begin{minipage}{0.3\textwidth}
                      \includegraphics[width=\textwidth]{Posterfig_After}
                      \caption{$Ro=-0.10$}
                    \end{minipage}
                    \hfill
                  \end{figure}

@@

            \begin{block}{\Large{Vortex lifetimes and cooling rate}}
              \justifying
              \begin{itemize}
              \item Vortices typically live longer with increasing cooling time:
              \end{itemize}
              \begin{figure}
                \centering
                \includegraphics[width=\textwidth,clip=true,trim=0cm
                  1.5cm 0cm 1cm]{lifetime_regime1}
              \end{figure}
              \vspace{-1cm}
              \begin{itemize}
              \item However, for very long $t_\mathrm{cool}$ vortex
                lifetimes shorten again:
              \end{itemize}
              \begin{figure}
                \centering
                \includegraphics[width=\textwidth,clip=true,trim=0cm
                  0cm 0cm 1cm]{lifetime_regime2}
              \end{figure}
              \vspace{-2cm}
              \begin{itemize}
              \item A cooling time of $\tilde{\beta}\sim 5$ maximizes
                vortex lifetime (nearly adiabatic)
              \end{itemize}
            \end{block}
            \vfill
            \begin{block}{{\Large Summary}}\justifying
              {\large{\bf
                  The stability of planetary gaps in gaseous
                  protoplanetary disks is affected by disk
                  thermodynamics. A slower cooling rate, which permits
                  more heating, tends to stabilize the system, but
                  also enables longer vortex lifetimes unless the
                  cooling time is extremely long. Our simulations
                  suggest there exists an optimal $t_\mathrm{cool}$
                  that maximizes vortex lifetimes, a result consistent
                  with recent isothermal simulations by Fu et
                  al. (2014,ApJ,788,L41) with variable disk aspect-ratio. 
                }
              }
              \vfill
            \end{block}
          }
          % ---------------------------------------------------------%
          % end the column
        \end{minipage}
      \end{beamercolorbox}
    \end{column}
    % ---------------------------------------------------------%
    % end the column
  \end{columns}
  \vskip1ex
  %\tiny\hfill\textcolor{ta2gray}{Created with \LaTeX \texttt{beamerposter}  \url{http://www-i6.informatik.rwth-aachen.de/~dreuw/latexbeamerposter.php}}
  %% \tiny\hfill{Created with \LaTeX \texttt{beamerposter}  \url{http://www-i6.informatik.rwth-aachen.de/~dreuw/latexbeamerposter.php} \hskip1em}
\end{frame}
\end{document}
